Loading libraries

```{r}
library(knitr)
library(ggplot2)
library(data.table)
```

Loading the data

```{r}
data <- read.csv("activity.csv", header=TRUE, sep=",")
str(data)
```

Process data

```{r}
data$date <- as.Date(data$date, format="%Y-%m-%d")
data$interval <- as.factor(data$interval)
str(data)
head(data, 10)
```

# Question 1 What is mean total number of steps taken per day?
## Calculate the total number of steps taken per day

```{r}
steps_per_day <- aggregate(steps ~ date, data=data, FUN=sum)
colnames(steps_per_day) <- c("date", "steps")
head(steps_per_day, 10)
```

## Make a histogram of the total number of steps taken each day

```{r}
ggplot(steps_per_day, aes(x = steps)) + 
        geom_histogram(fill = "blueviolet", binwidth = 1000) + 
        labs(title = "Histogram - Steps Taken Per Day", x = "Steps Per Day", y = "Frequency")
```

## Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean_steps_per_day <- mean(steps_per_day$steps)
mean_steps_per_day
median_steps_per_day <- median(steps_per_day$steps)
median_steps_per_day
```

# Question 2 What is the average daily activity pattern?

## Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps_per_interval <- aggregate(steps ~ interval, data = data, FUN = mean, na.rm = TRUE)
steps_per_interval$interval <- as.integer(levels(steps_per_interval$interval)[steps_per_interval$interval])
colnames(steps_per_interval) <- c("interval", "steps")
head(steps_per_interval, 10)
```

## Plot

```{r}
ggplot(steps_per_interval, aes(x = interval, y = steps)) + 
        geom_line(col = "cornflowerblue", size = 1) + 
        labs(title = "Average Daily Activity Pattern", x = "Interval", y = "Steps")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_interval <- steps_per_interval[which.max(steps_per_interval$steps),]
max_interval
```

# Question 3 Imputing missing values
## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missing_values <- sum(is.na(data$steps))
missing_values
```

## Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newdata <- data
index_of_na <- which(is.na(newdata$steps))
for (i in index_of_na) {
        newdata$steps[i] <- with(steps_per_interval, steps[interval = newdata$interval[i]])
}
head(newdata, 10)
new_missing_values <- sum(is.na(newdata$steps))
new_missing_values
```

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
new_steps_per_day <- aggregate(steps ~ date, data = newdata, FUN=sum)
colnames(new_steps_per_day) <- c("date", "steps")
ggplot(new_steps_per_day, aes(x = steps)) + 
        geom_histogram(fill = "forestgreen", binwidth = 1000) + 
        labs(title = "Histogram - Steps Taken Per Day", x = "Steps Per Day", y = "Frequency")

new_mean_steps_per_day <- mean(new_steps_per_day$steps)
new_mean_steps_per_day
new_median_steps_per_day <- median(new_steps_per_day$steps)
new_median_steps_per_day
```

# Question 4 Are there differences in activity patterns between weekdays and weekends?
## Create a new factor variable in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
dt <- data.table(newdata)
dt[, weekday := ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")]
dt$weekday <- as.factor(dt$weekday)
dt$interval <- as.integer(levels(dt$interval)[dt$interval])
head(dt, 10)
```

## Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
steps_per_weekday <- aggregate(steps ~ interval+weekday, data = dt, FUN = mean)
ggplot(steps_per_weekday, aes(x = interval, y = steps)) + 
        geom_line(col = "blue", size = 1) + 
        facet_wrap(~ weekday, nrow=2, ncol=1) + 
        labs(x = "Interval", y = "Number of Steps")
```
